Thread Pool in GeckoDB/BOLSTER - Integrating a Thread Pool into a high performance Database System \\

### 2018-06-21 Presentations/Paper/paper_thread_pool.tex
added some text about dynamic resizing reduced some word repetitions

Threadsafe access to memory can be achieved by different approaches. Firstly, a mutex thread can manage the access for multiple other threads. Secondly, atomic operations ensure the safe access from different threads to the same memory. We use both approaches for different parts of the thread pool. Since using a mutex thread can decrease access performance, we decided to execute atomic operations on the memory containing the thread state informations. For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation. We use a mutex thread to synchronize the enqueueing of tasks into the priority queue.

@@

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, the state has to be updated. If these steps are implemented using an if statement and an update function, the complete update procedure is not atomic. Consequently, another thread can update the state between the check of the current state and the update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.

>>>>

Thread safe access to memory can be achieved by different approaches. Firstly, a mutex thread can manage the access for multiple other threads. Secondly, atomic operations ensure the safe access from different threads to the same memory. We use both approaches for different parts of the thread pool. Since using a mutex thread can decrease access performance, we decided to execute atomic operations on memory containing the thread state informations. For example, the function \emph{atomic\_compare\_exchange\_strong} performs a compare and, if the result is true, an exchange of the memory in one atomic operation. We use a mutex thread to synchronize the enqueueing of tasks into the priority queue.

@@

\subsection{Managing Thread States}
As mentioned before, accessing the same memory location from different thread may lead to race conditions, which means the program behaviour depends on the accessing order of the threads. For example, while one thread updates data at a specific memory location another threads writes into the data, which results to inconsistency. In Section 2, we presented two approaches to handle multithreaded access to the same memory location, mutex threads and atomic operations and we state where those approaches are used in our implementation. In this subsection, we focus on atomic operations to manage thread states. We consider the following steps for a thread to update its state. At first, the current state has to be checked and, depending on the current state, it has to be updated. If these steps are implemented using an if statement and an update function, the complete procedure is not atomic. Consequently, another thread can update the state between check of the current state and update. To avoid this, we use the function \emph{atomic\_compare\_exchange\_strong} \cite{atomicstrong} of the stdatomic library. This function combines the check and update in one atomic procedure. We implemented every thread state access with this function to ensure atomic thread state updates. \\
 This approach enables efficient dynamic and lazy resizing of the thread pool. When the number of threads is reduced, the now redundant threads are marked to finish after completing their current task. If later on the number is increased again, still running but marked threads can be enabled again ensuring that few threads have to be created and that the number of running threads will not temporarily exceed the requested size. \\
 After presenting the enqueueing, processing, and waiting for tasks, we show the second component of our thread pool, the performance monitoring.
